### 2023-04-18 main.tex
ll
i

\newcommand{\KL}{\mathcal{K}}
\newcommand{\KR}{\hat{\mathcal{K}}}
\newcommand{\oa}{\mathbf{a}}
\newcommand{\dd}{\mathcal{D}_{\widehat{x}}}
\newcommand{\oad}{\mathbf{a}^{\dagger}}

@@

This means that on the edge $(x,y)\in \mathcal{E}$ at rate $\omega_{x,y}n_{a}^{x}n_{b}^{y}$ a particle of type $a$ at site $x$ is exchanged with a particle of type $b$ at site $y$; furthermore, on each site $x\in V$ at rate $\Gamma_{x}\alpha_{a}^{x}n_{b}^{x}$ a particle of type $b$ is replaced with a particle of type $a$. 

@@

\section{Duality}\label{sectionDuality}
\subsection{Definition}
Consider two Markov processes $(\eta_{t})_{t\geq 0}$ defined on a state space $\Omega$ and $(\xi_{t})_{t\geq 0}$ defined on a state space $\widetilde{\Omega}$. We say that they are dual, with respect to a duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$, if $\forall \eta\in\Omega$, $\forall \xi\in\widetilde{\Omega}$ and $\forall t> 0$ we have 
\begin{equation}
    \mathbb{E}_{\eta}\left[D(\eta_{t},\xi)\right]=\mathbb{E}_{\xi}\left[D(\eta,\xi_{t})\right]
\end{equation}
where $\mathbb{E}_{\eta}$ denotes the expectation with respect the law of the Markov process $(\eta_{t})_{t\geq 0}$ initialized with the particle configuration $\eta$, whereas $\mathbb{E}_{\xi}$ denotes the expectation with respect to the law of the Markov process $(\xi_{t})_{t\geq 0}$ initialized with the particle configuration $\xi$.
The duality definition can also be formulated as a relation between the generators. Call $\mathcal{L}$ the generator of $(\eta_{t})_{t\geq0}$ and $\widetilde{\mathcal{L}}$ the generator of $(\xi_{t})_{t\geq 0}$, then we say that these two processes are dual with respect to the duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$ if $\forall \eta\in\Omega$ and $\forall \xi\in\widetilde{\Omega}$
\begin{equation}\label{dualityRelationGenerator}
    \left(\mathcal{L}D(\cdot,\xi)\right)(\eta)=\left(\widetilde{\mathcal{L}}D(\eta,\cdot)\right)(\xi)
\end{equation}
In the specific case where $\mathcal{L}=\widetilde{\mathcal{L}}$ we say that the process is self-dual.
\newline
\newline
\textbf{Reamrk}:
when the state spaces of the dual processes is finite, the generators and the duality function are matrices with elements $\mathcal{L}(\eta,\eta^{'})$, $\widetilde{\mathcal{L}}(\xi,\xi^{'})$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta^{'}\in\Omega$ and $\xi,\xi^{'}\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 
\begin{equation}
    \sum_{\eta^{'}\in\,\Omega}\mathcal{L}(\eta,\eta^{'})D(\eta^{'},\xi)=\sum_{\xi^{'}\in\, \widetilde{\Omega}}\widetilde{\mathcal{L}}(\xi,\xi^{'})D(\eta,\xi^{'})
\end{equation}
that can be read as
\begin{equation}\label{dualityIntertwines}
    \mathcal{L}D=D\widetilde{\mathcal{L}}^{\,T}
\end{equation}
where the superscript $T$ denotes the matrix transposition. Therefore, the duality relation \eqref{dualityIntertwines} is an intertwining between two linear operators $\mathcal{L}$ and $\widetilde{\mathcal{L}}$. Working with the Hamiltonian operators the duality relation \eqref{dualityIntertwines} reads 
\begin{equation}\label{DualityRelation}
    H^{T}D=D\widetilde{H}
\end{equation}
\subsection{Duality for the multi-species stirring process}\label{statementDualitySubsection}
In this section we formulate duality for the multi-species stirring process $(\mathbf{n}(t))_{t\geq 0}$, defined by the generator \eqref{Generator}.
The dual  process $(\bm{\xi}(t))_{t\geq 0}$ takes values in the enlarged state space
\begin{equation}\label{dualStateSpace}
    \widetilde{\Omega}= \bigtimes_{x\in V} \widetilde{\Omega}_{x}\ = \bigtimes_{x\in V} (\Omega_{x}\times \mathbb{N}_{0}^{N-1})
\end{equation}
To each site $x\in V$ we associate an ``extra site'', denoted $\widehat{x}$,
where dual particles will accumulate in the course of time. 
We write the configurations $\bm{\xi} \in \widetilde\Omega$  as
\begin{equation}
    \bm{\xi}=\left(r_{1}^{x},\ldots,r_{N}^{x},m_{2}^{\widehat{x}},\ldots,m_{N}^{\widehat{x}}\right)_{x\in V}
\end{equation}
where the component $r_{a}^{x}$ is interpreted as the number of dual particles of type $a\in \{1,\ldots,N\}$ at site $x$, 
and the component $m_{a}^{\widehat{x}}$  gives the number of dual particles of type $a\in \{2,\ldots,N\}$ at 
the extra-site $\widehat{x}$ connected to $x\in V$.

The duality function is given by
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(\nu -\sum_{a=2}^{N}r_{a}^{x})!}{\nu!}\prod_{a=2}^{N}\frac{n_{a}^{x}!}{(n_{a}^{x}-r_{a}^{x})!}\left(\rho_{a}^{x}\right)^{m_{a}^{\widehat{x}}}\,\right)
\end{equation}
where we denote the \textit{density} of the species $a\in \{2,\ldots,N\}$ imposed by the reservoir at site $x\in V$  by 
\begin{equation}
\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{\sum_{i=1}^{N}\alpha_{i}^{x}}
\end{equation}
The dual process is defined by its generator, which reads
 \begin{equation}\label{DualGenerator}
    \widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
\end{equation}
where 
$\mathcal{L}_{x,y}$ is defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
    \widetilde{\mathcal{L}}_{x}f(\bm{\xi})=\sum_{i=1}^{N}\alpha_{i}^{x}\sum_{a=2}^{N}r_{a}^{x}\left(f(\bm{\xi}-\delta_{a}^{x}+\delta_{1}^{x}+\delta_{a}^{\widehat{x}})-f(\bm{\xi})\right)
\end{equation}
\newline
Therefore the dual dynamics is described as follows. On one hand, the edge part $\mathcal{L}_{x,y}$ of the dual generator gives rise to  the multi-species stirring dynamics on the graph. On the other hand, the site
part $\widetilde{\mathcal{L}}_{x}$ of the dual generator replaces a particle of any type $a\in\{2,\ldots,N\}$ at site $x$ with a particle of type $1$ and creates a particle of the same type $a$ at the extra-site $\widehat{x}$. This last transition is performed with rate $\sum_{i=1}^{N}\alpha_{i}^{x}r_{a}^{x}$. This means that eventually the dual process voids the graph, putting all the dual particles of species $\{2,\ldots,N\}$ in the extra-sites. In other words the extra-sites play the role of absorbing boundaries. 
\newline \newline
\textbf{Remark}: in the reversible situation, i.e. when $\forall x\in V$ $\rho_{a}^{x}=\rho_{a}$, the expectation  of the duality function  $D(\bm{n},\bm{\xi})$ with   $\bm{n}$ distributed as  $\mu_{rev} = \bigtimes_{x\in V}\text{Multinomial}\left(\twoj, \rho_{1},\ldots,\rho_{N}\right)$ is
\begin{equation}
\mathbb{E}_{\mu^{rev}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{a=2}^{N}\left(\rho_{a}\right)^{\sum_{x\in V}r_{a}^{x}+\sum_{x\in V}m_{a}^{\widehat{x}}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}
\end{equation}

\subsection{Proof of duality}
To prove duality between $(\bm{n}(t))_{t\geq 0}$ and $(\bm{\xi}(t))_{t\geq 0}$ we  show that \eqref{dualityIntertwines} is fulfilled.  To show this, we will use the Hamiltonians (linked with the generators by \eqref{Hamiltonian-Generator}) and their Lie algebraic description. Indeed, in this formalism the proof reduces to finding symmetries and exponential transformation of generators of the Lie algebra.\\
To each configuration $\bm{\xi}$ of the configuration space  \eqref{dualStateSpace} of the dual process we associate the vector
    |\bm{\xi}\rangle=\bigotimes_{x\in V}\left(|r_{1}^{x},\ldots,r_{N}^{x}\rangle\otimes |m_{2}^{\widehat{x}},\ldots,m_{N}^{\widehat{x}}\rangle\right)
The Hamiltonian of the dual process reads
\begin{equation}\label{DualHamiltonian}
    \widetilde{H}=\sum_{x,y\in \mathcal{E}}\omega_{x,y}\mathcal{H}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{H}_{x}
where $\mathcal{H}_{x,y}$ is the one defined in \eqref{edgeHamiltonian}, while 
\begin{equation}\label{siteDualHamiltonian}
    \widetilde{H}_{x}=\sum_{i=1}^{N}\alpha_{i}^{x}\sum_{a=2}^{N}\left((a^{\dagger})_{a}^{\widehat{x}}\,E_{1a}^{x}-E_{aa}^{x}\right)
Here we introduced bosonic creation operator $a^{\dagger}$ acting as $a^{\dagger}|q\rangle=|q+1\rangle$ on a generic vector $|q\rangle$ with $q\in \mathbb{N}_{0}$, so that in \eqref{siteDualHamiltonian} 
$(a^{\dagger})_{a}^{\widehat{x}}$ denotes a copy of $a^{\dagger}$ acting on the extra-site $\widehat{x}$ and on the species $a\in\{2,\ldots,N\}$. \\
We will show below that the Hamiltonians \eqref{OriginalHamiltonian} and \eqref{DualHamiltonian} are dual in the sense of \eqref{DualityRelation}. The duality matrix $D$ is described as 
\begin{equation}\label{dualityMatrix}
    D=\prod_{x\in V}d_{x}\otimes \dd
\begin{equation}\label{bulkElementDualityMatrix}
d_{x}=R_{x}\exp{(E^{x})}
\begin{equation}\label{EquationEx}
E^{x}=\sum_{a=2}^{N}E_{a1}^{x}
\begin{equation}\label{Revmatrix}
    R_{x}=\sum_{n^{x}\in\Omega_{x}}\frac{\prod_{a=1}^{N}n_{a}^{x}!}{\nu!}|n_{1}^{x},\ldots,n_{N}^{x}\rangle\langle n_{1}^{x},\ldots,n_{N}^{x}|
\begin{equation}\label{dualityMatrix2}
\dd=\sum_{m_{2}^{\widehat{x}},\ldots,m_{N}^{\widehat{x}}=0}^{\infty}\prod_{a=2}^{N}\left(\rho_{a}^{x}\right)^{m_{a}^{\widehat{x}}}\langle m_{2}^{\widehat{x}},\ldots,m_{N}^{\widehat{x}}|
The matrix $R_{x}$ is diagonal. Its elements are related to the inverse of the weights of the reversible measure \eqref{reversibleMeasure}. In particular, to obtain these elements, we have considered the weights of \eqref{reversibleMeasure} when all the parameters $p_{i}=\frac{1}{N}$. Then, the constant $\left(\frac{1}{N}\right)^{\nu}$ has been neglected, since it does not change the duality relation. This $R_{x}$ is called the ``cheap'' duality matrix (see \cite{giardina2009duality}). \\
Since \eqref{dualityMatrix} is product over sites, proving \eqref{DualityRelation} is equivalent to showing that 
\begin{equation}\label{edgeDualRealtion}
    \mathcal{H}_{x,y}^{T}D=D\mathcal{H}_{x,y}\qquad \forall (x,y)\in \mathcal{E}
\begin{equation}\label{siteDualRelation}
    H_{x}^{T}D=D\widetilde{H}_{x}\qquad \forall x\in V.
We perform the duality proof in three steps: first we will show that matrix \eqref{dualityMatrix} has elements \eqref{dualityElements}; second we will prove \eqref{edgeDualRealtion}; finally we will show\eqref{siteDualRelation}. 
As a preliminary result we show the following property 
\begin{equation}\label{transpositionPropertyR}
(E_{ba}^{x})^{T}=R_{x}E_{ab}^{x}R_{x}^{-1}\qquad \forall x\in V
R_{x}E_{ab}^{x}R_{x}^{-1}=&\sum_{r^{x}\in\Omega_{x}}\frac{r_{1}^{x}!\ldots r_{N}!}{\nu!}|r_{1}^{x},\ldots,r_{N}^{x}\rangle \langle r_{1}^{x},\ldots, r_{N}^{x}|

@@

	r_{a}^{x}|r_{1}^{x},\ldots,r_{N}^{x}\rangle \langle r_{1}^{x},\ldots,r_{a}^{x}-1,\ldots,r_{b}^{x}-1,\ldots,r_{N}^{x}|

@@

	 &\langle n^{x}|d_{x}\otimes \dd|\xi^{x}\rangle\\&=\langle n_{1}^{x},\ldots,n_{N}^{x}| (\exp{(E_{12}^{x}+\ldots+E_{1N}^{x}}))^{T}R_{x}\otimes\sum_{q_{2}^{\widehat{x}},\ldots,q_{N}^{\widehat{x}}=0}^{\infty}\prod_{a=2}^{N}\left(\rho_{a}^{x}\right)^{q_{a}^{\widehat{x}}}\langle q_{2}^{\widehat{x}},\ldots,q_{N}^{\widehat{x}}|
	 \\&|r_{1}^{x},\ldots,r_{N}^{x}\rangle_{x} \otimes |m_{2}^{\widehat{x}},\ldots,m_{N}^{\widehat{x}}\rangle

@@

On one hand, on the extra-site $\widehat{x}$ we have 
\sum_{q_{2}^{\widehat{x}},\ldots,q_{N}^{\widehat{x}}=0}^{\infty}\prod_{a=2}^{N}\left(\rho_{a}^{x}\right)^{q_{a}^{\widehat{x}}}\langle q_{2}^{\widehat{x}},\ldots,q_{N}^{\widehat{x}}|m_{2}^{\widehat{x}},\ldots,m_{N}^{\widehat{x}}\rangle=\prod_{a=2}^{N}\left(\rho_{a}^{x}\right)^{m_{a}^{\widehat{x}}}

@@

&\langle n_{1}^{x},\ldots,n_{N}^{x}|(\exp{(E_{12}^{x}+\ldots+E_{1N}^{x})})^{T}R_{x}|r_{1}^{x},\ldots,r_{N}^{x}\rangle\\&= \langle  n_{1}^{x},\ldots,n_{N}^{x}|\left(\sum_{k_{2}=0}^{\infty}\frac{\left\{\left(E_{12}^{x}\right)^{T}\right\}^{k_{2}}}{k_{2}!}\ldots\sum_{k_{N}=0}^{\infty}\frac{\left\{\left(E_{1N}^{x}\right)^{T}\right\}^{k_{N}}}{k_{N}!}\sum_{s\in\Omega_{x}}\frac{s_{1}^{x}!\ldots s_{N}^{x}!}{\nu!}|s_{1}^{x},\ldots,s_{N}^{x}\rangle\langle s_{1}^{x},\ldots,s_{N}^{x}|\right)

@@

\hspace{2.cm}\qquad|r_{1}^{x},\ldots,r_{N}^{x}\rangle

@@

\hspace{2.cm}\prod_{a=2}^{N}\left((n_{a}^{x}(n_{a}^{x}-1)\ldots (n_{a}^{x}-k_{a}+1)\right)\frac{1}{k_{2}!,\ldots,k_{N}!}\frac{r_{1}^{x}!\ldots r_{N}^{x}!}{\nu!}|r_{1}^{x},\ldots,r_{N}^{x}\rangle

@@

\hspace{2.cm}\qquad\frac{1}{k_{2}!,\ldots,k_{N}!}\frac{r_{1}^{x}!\ldots r_{N}^{x}!}{\nu!}|r_{1}^{x},\ldots,r_{N}^{x}\rangle
\frac{(\nu-\sum_{a=2}^{N}r_{a}^{x})}{\nu!}\prod_{a=2}^{N}\frac{n_{a}^{x}!}{(n_{a}^{x}-r_{a}^{x})!}
where in the last equality we applied the orthogonality relations \eqref{ortho} and the fact that $r_{1}^{x}=\nu-\sum_{a=2}^{N}r_{a}^{x}$. Finally, by taking the product over $x\in V$ \eqref{proofDualityElements} is proved.

@@

    \sum_{a=2}^{N}\alpha_{a}^{x}E_{1a}^{x}-\sum_{i=1}^{N}\alpha_{i}^{x}\sum_{b=2}^{N}E_{bb}^{x}=
    \sum_{a=2}^{N}\left(\alpha_{a}^{x}E_{1a}^{x}-\sum_{i=1}^{N}\alpha_{i}^{x}E_{aa}^{x}\right)

@@

		\langle n^{y_{1}}_{s_{1}}\cdots n^{y_{m}}_{s_{m}}\rangle_{\text{stat}}=\sum_{c_{1},\ldots,c_{m}=0}^{1}

>>>>

\newcommand{\KL}{\mathcal{K}}
\newcommand{\KR}{\hat{\mathcal{K}}}
\newcommand{\oa}{\mathbf{a}}
\newcommand{\dd}{\mathcal{D}_{u(x)}}
\newcommand{\oad}{\mathbf{a}^{\dagger}}

@@

This means that on the edge $(x,y)\in \mathcal{E}$ at rate $\omega_{x,y}n_{a}^{x}n_{b}^{y}$ a particle of type $a$ at site $x$ is exchanged with a particle of type $b$ at site $y$; furthermore, on each site $x\in V$ at rate $\Gamma_{x}\alpha_{a}^{x}n_{b}^{x}$ a particle of type $b$ is replaced with a particle of type $a$. For the sake of notation we denote $|\alpha^{x}|=\sum_{a=1}^{N}\alpha_{a}^{x}$. 

@@

\section{Duality}\label{sectionDuality}
\subsection{Definition}
Consider two Markov processes $(\eta_{t})_{t\geq 0}$ defined on a state space $\Omega$ and $(\xi_{t})_{t\geq 0}$ defined on a state space $\widetilde{\Omega}$. We say that they are dual, with respect to a duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$, if $\forall \eta\in\Omega$, $\forall \xi\in\widetilde{\Omega}$ and $\forall t> 0$ we have 
\begin{equation}
    \mathbb{E}_{\eta}\left[D(\eta_{t},\xi)\right]=\mathbb{E}_{\xi}\left[D(\eta,\xi_{t})\right]
\end{equation}
where $\mathbb{E}_{\eta}$ denotes the expectation with respect the law of the Markov process $(\eta_{t})_{t\geq 0}$ initialized with the particle configuration $\eta$, whereas $\mathbb{E}_{\xi}$ denotes the expectation with respect to the law of the Markov process $(\xi_{t})_{t\geq 0}$ initialized with the particle configuration $\xi$.
The duality definition can also be formulated as a relation between the generators. Call $\mathcal{L}$ the generator of $(\eta_{t})_{t\geq0}$ and $\widetilde{\mathcal{L}}$ the generator of $(\xi_{t})_{t\geq 0}$, then we say that these two processes are dual with respect to the duality function $D:\Omega\times \widetilde{\Omega}\to \mathbb{R}$ if $\forall \eta\in\Omega$ and $\forall \xi\in\widetilde{\Omega}$
\begin{equation}\label{dualityRelationGenerator}
    \left(\mathcal{L}D(\cdot,\xi)\right)(\eta)=\left(\widetilde{\mathcal{L}}D(\eta,\cdot)\right)(\xi)
\end{equation}
In the specific case where $\mathcal{L}=\widetilde{\mathcal{L}}$ we say that the process is self-dual.
\newline
\newline
\textbf{Reamrk}:
when the state spaces of the dual processes is finite, the generators and the duality function are matrices with elements $\mathcal{L}(\eta,\eta^{'})$, $\widetilde{\mathcal{L}}(\xi,\xi^{'})$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta^{'}\in\Omega$ and $\xi,\xi^{'}\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 
\begin{equation}
    \sum_{\eta^{'}\in\,\Omega}\mathcal{L}(\eta,\eta^{'})D(\eta^{'},\xi)=\sum_{\xi^{'}\in\, \widetilde{\Omega}}\widetilde{\mathcal{L}}(\xi,\xi^{'})D(\eta,\xi^{'})
\end{equation}
that can be read as
\begin{equation}\label{dualityIntertwines}
    \mathcal{L}D=D\widetilde{\mathcal{L}}^{\,T}
\end{equation}
where the superscript $T$ denotes the matrix transposition. Therefore, the duality relation \eqref{dualityIntertwines} is an intertwining between two linear operators $\mathcal{L}$ and $\widetilde{\mathcal{L}}$. Working with the Hamiltonian operators the duality relation \eqref{dualityIntertwines} reads 
\begin{equation}\label{DualityRelation}
    H^{T}D=D\widetilde{H}
\end{equation}
\subsection{Duality for the multi-species stirring process}\label{statementDualitySubsection}
In this section we formulate duality for the multi-species stirring process $(\mathbf{n}(t))_{t\geq 0}$, defined by the generator \eqref{Generator}.
The dual  process $(\bm{\xi}(t))_{t\geq 0}$ takes values in the enlarged state space
\begin{equation}\label{dualStateSpace}
    \widetilde{\Omega}= \bigtimes_{x\in V} \widetilde{\Omega}_{x}\ = \bigtimes_{x\in V} (\Omega_{x}\times \mathbb{N}_{0}^{N-1})
\end{equation}
To each site $x\in V$ we associate an ``extra site'', denoted $u(x)$,
where dual particles will accumulate in the course of time. 
We write the configurations $\bm{\xi} \in \widetilde\Omega$  as
\begin{equation}
    \bm{\xi}=\left(\xi_{1}^{x},\ldots,\xi_{N}^{x},\xi_{2}^{u(x)},\ldots,\xi_{N}^{u(x)}\right)_{x\in V}
\end{equation}
where the component $\xi_{a}^{x}$ is interpreted as the number of dual particles of type $a\in \{1,\ldots,N\}$ at site $x$, 
and the component $m_{a}^{u(x)}$  gives the number of dual particles of type $a\in \{2,\ldots,N\}$ at 
the extra-site $u(x)$ connected to $x\in V$.

The duality function is given by
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(\nu -\sum_{a=2}^{N}\xi_{a}^{x})!}{\nu!}\prod_{a=2}^{N}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\left(\rho_{a}^{x}\right)^{m_{a}^{u(x)}}\,\right)
\end{equation}
where we denote the \textit{density} of the species $a\in \{2,\ldots,N\}$ imposed by the reservoir at site $x\in V$  by 
\begin{equation}
\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}
\end{equation}
The dual process is defined by its generator, which reads
 \begin{equation}\label{DualGenerator}
    \widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
\end{equation}
where 
$\mathcal{L}_{x,y}$ is defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
    \widetilde{\mathcal{L}}_{x}f(\bm{\xi})=|\alpha^{x}|\sum_{a=2}^{N}\xi_{a}^{x}\left(f(\bm{\xi}-\delta_{a}^{x}+\delta_{1}^{x}+\delta_{a}^{u(x)})-f(\bm{\xi})\right)
\end{equation}
\newline
Therefore the dual dynamics is described as follows. On one hand, the edge part $\mathcal{L}_{x,y}$ of the dual generator gives rise to  the multi-species stirring dynamics on the graph. On the other hand, the site
part $\widetilde{\mathcal{L}}_{x}$ of the dual generator replaces a particle of any type $a\in\{2,\ldots,N\}$ at site $x$ with a particle of type $1$ and creates a particle of the same type $a$ at the extra-site $u(x)$. This last transition is performed with rate $|\alpha^{x}|\xi_{a}^{x}$. This means that eventually the dual process voids the graph, putting all the dual particles of species $\{2,\ldots,N\}$ in the extra-sites. In other words the extra-sites play the role of absorbing boundaries. 
\newline \newline
\textbf{Remark}: in the reversible situation, i.e. when $\forall x\in V$ $\rho_{a}^{x}=\rho_{a}$, the expectation  of the duality function  $D(\bm{n},\bm{\xi})$ with   $\bm{n}$ distributed as  $\mu_{rev} = \bigtimes_{x\in V}\text{Multinomial}\left(\twoj, \rho_{1},\ldots,\rho_{N}\right)$ is
\begin{equation}
\mathbb{E}_{\mu^{rev}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{a=2}^{N}\left(\rho_{a}\right)^{\sum_{x\in V}\xi_{a}^{x}+\sum_{x\in V}m_{a}^{u(x)}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}
\end{equation}

\subsection{Proof of duality}
To prove duality between $(\bm{n}(t))_{t\geq 0}$ and $(\bm{\xi}(t))_{t\geq 0}$ we  show that \eqref{dualityIntertwines} is fulfilled.  To show this, we will use the Hamiltonians (linked with the generators by \eqref{Hamiltonian-Generator}) and their Lie algebraic description. Indeed, in this formalism the proof reduces to finding symmetries and exponential transformation of generators of the Lie algebra.\\
To each configuration $\bm{\xi}$ of the configuration space  \eqref{dualStateSpace} of the dual process we associate the vector
\begin{equation}
    |\bm{\xi}\rangle=\bigotimes_{x\in V}\left(|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\otimes |\xi_{2}^{u(x)},\ldots,\xi_{N}^{u(x)}\rangle\right)
\end{equation}
The Hamiltonian of the dual process reads
\begin{equation}\label{DualHamiltonian}
    \widetilde{H}=\sum_{x,y\in \mathcal{E}}\omega_{x,y}\mathcal{H}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{H}_{x}
\end{equation}
where $\mathcal{H}_{x,y}$ is the one defined in \eqref{edgeHamiltonian}, while 
\begin{equation}\label{siteDualHamiltonian}
    \widetilde{H}_{x}=|\alpha^{x}|\sum_{a=2}^{N}\left((a^{\dagger})_{a}^{u(x)}\,E_{1a}^{x}-E_{aa}^{x}\right)
\end{equation}
Here we introduced bosonic creation operator $a^{\dagger}$ acting as $a^{\dagger}|q\rangle=|q+1\rangle$ on a generic vector $|q\rangle$ with $q\in \mathbb{N}_{0}$, so that in \eqref{siteDualHamiltonian} 
$(a^{\dagger})_{a}^{u(x)}$ denotes a copy of $a^{\dagger}$ acting on the extra-site $u(x)$ and on the species $a\in\{2,\ldots,N\}$. \\
We will show below that the Hamiltonians \eqref{OriginalHamiltonian} and \eqref{DualHamiltonian} are dual in the sense of \eqref{DualityRelation}. The duality matrix $D$ is described as 
\begin{equation}\label{dualityMatrix}
    D=\prod_{x\in V}d_{x}\otimes \dd
\end{equation}
where
\begin{equation}\label{bulkElementDualityMatrix}
d_{x}=R_{x}\exp{(E^{x})}
\end{equation}
with 
\begin{equation}\label{EquationEx}
E^{x}=\sum_{a=2}^{N}E_{a1}^{x}
\end{equation}
and
\begin{equation}\label{Revmatrix}
    R_{x}=\sum_{n^{x}\in\Omega_{x}}\frac{\prod_{a=1}^{N}n_{a}^{x}!}{\nu!}|n_{1}^{x},\ldots,n_{N}^{x}\rangle\langle n_{1}^{x},\ldots,n_{N}^{x}|
\end{equation}
and where 
\begin{equation}\label{dualityMatrix2}
\dd=\sum_{\xi_{2}^{u(x)},\ldots,\xi_{N}^{u(x)}=0}^{\infty}\prod_{a=2}^{N}\left(\rho_{a}^{x}\right)^{m_{a}^{u(x)}}\langle \xi_{2}^{u(x)},\ldots,\xi_{N}^{u(x)}|
\end{equation}
The matrix $R_{x}$ is diagonal. Its elements are related to the inverse of the weights of the reversible measure \eqref{reversibleMeasure}. In particular, to obtain these elements, we have considered the weights of \eqref{reversibleMeasure} when all the parameters $p_{i}=\frac{1}{N}$. Then, the constant $\left(\frac{1}{N}\right)^{\nu}$ has been neglected, since it does not change the duality relation. This $R_{x}$ is called the ``cheap'' duality matrix (see \cite{giardina2009duality}). \\
Since \eqref{dualityMatrix} is product over sites, proving \eqref{DualityRelation} is equivalent to showing that 
\begin{equation}\label{edgeDualRealtion}
    \mathcal{H}_{x,y}^{T}D=D\mathcal{H}_{x,y}\qquad \forall (x,y)\in \mathcal{E}
\end{equation}
and 
\begin{equation}\label{siteDualRelation}
    H_{x}^{T}D=D\widetilde{H}_{x}\qquad \forall x\in V.
\end{equation}
We perform the duality proof in three steps: first we will show that matrix \eqref{dualityMatrix} has elements \eqref{dualityElements}; second we will prove \eqref{edgeDualRealtion}; finally we will show\eqref{siteDualRelation}. 
As a preliminary result we note that 
\begin{equation}\label{transpositionPropertyR}
(E_{ba}^{x})^{T}=R_{x}E_{ab}^{x}R_{x}^{-1}\qquad \forall x\in V
\end{equation}
that follows immediately from the definition of $E_{ab}$. 
\begin{comment}
R_{x}E_{ab}^{x}R_{x}^{-1}=&\sum_{r^{x}\in\Omega_{x}}\frac{\xi_{1}^{x}!\ldots r_{N}!}{\nu!}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle \langle \xi_{1}^{x},\ldots, \xi_{N}^{x}|

@@

	\xi_{a}^{x}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle \langle \xi_{1}^{x},\ldots,\xi_{a}^{x}-1,\ldots,r_{b}^{x}-1,\ldots,\xi_{N}^{x}|

@@

	 &\langle n^{x}|d_{x}\otimes \dd|\xi^{x}\rangle\\&=\langle n_{1}^{x},\ldots,n_{N}^{x}| (\exp{(E_{12}^{x}+\ldots+E_{1N}^{x}}))^{T}R_{x}\otimes\sum_{q_{2}^{u(x)},\ldots,q_{N}^{u(x)}=0}^{\infty}\prod_{a=2}^{N}\left(\rho_{a}^{x}\right)^{q_{a}^{u(x)}}\langle q_{2}^{u(x)},\ldots,q_{N}^{u(x)}|
	 \\&|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle \otimes |\xi_{2}^{u(x)},\ldots,\xi_{N}^{u(x)}\rangle

@@

On one hand, on the extra-site $u(x)$ we have 
\sum_{q_{2}^{u(x)},\ldots,q_{N}^{u(x)}=0}^{\infty}\prod_{a=2}^{N}\left(\rho_{a}^{x}\right)^{q_{a}^{u(x)}}\langle q_{2}^{u(x)},\ldots,q_{N}^{u(x)}|\xi_{2}^{u(x)},\ldots,\xi_{N}^{u(x)}\rangle=\prod_{a=2}^{N}\left(\rho_{a}^{x}\right)^{m_{a}^{u(x)}}

@@

&\langle n_{1}^{x},\ldots,n_{N}^{x}|(\exp{(E_{12}^{x}+\ldots+E_{1N}^{x})})^{T}R_{x}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\\&= \langle  n_{1}^{x},\ldots,n_{N}^{x}|\left(\sum_{k_{2}=0}^{\infty}\frac{\left\{\left(E_{12}^{x}\right)^{T}\right\}^{k_{2}}}{k_{2}!}\ldots\sum_{k_{N}=0}^{\infty}\frac{\left\{\left(E_{1N}^{x}\right)^{T}\right\}^{k_{N}}}{k_{N}!}\sum_{s\in\Omega_{x}}\frac{s_{1}^{x}!\ldots s_{N}^{x}!}{\nu!}|s_{1}^{x},\ldots,s_{N}^{x}\rangle\langle s_{1}^{x},\ldots,s_{N}^{x}|\right)

@@

\hspace{2.cm}\qquad|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle

@@

\hspace{2.cm}\prod_{a=2}^{N}\left((n_{a}^{x}(n_{a}^{x}-1)\ldots (n_{a}^{x}-k_{a}+1)\right)\frac{1}{k_{2}!,\ldots,k_{N}!}\frac{\xi_{1}^{x}!\ldots \xi_{N}^{x}!}{\nu!}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle

@@

\hspace{2.cm}\qquad\frac{1}{k_{2}!,\ldots,k_{N}!}\frac{\xi_{1}^{x}!\ldots \xi_{N}^{x}!}{\nu!}|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle
\frac{(\nu-\sum_{a=2}^{N}\xi_{a}^{x})}{\nu!}\prod_{a=2}^{N}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}
where in the last equality we applied the orthogonality relations \eqref{ortho} and the fact that $\xi_{1}^{x}=\nu-\sum_{a=2}^{N}\xi_{a}^{x}$. Finally, by taking the product over $x\in V$ \eqref{proofDualityElements} is proved.

@@

    \sum_{a=2}^{N}\alpha_{a}^{x}E_{1a}^{x}-|\alpha^{x}|\sum_{b=2}^{N}E_{bb}^{x}=
    \sum_{a=2}^{N}\left(\alpha_{a}^{x}E_{1a}^{x}-|\alpha^{x}|E_{aa}^{x}\right)

@@

		\langle n^{x_{1}}_{a_{1}}\cdots n^{x_{m}}_{a_{m}}\rangle_{\text{stat}}=\sum_{c_{1},\ldots,c_{m}=0}^{1}
